---
title: "Association Rules Project"
output:
  html_notebook: default
  word_document: default
---

# 1.Load libraries
```{r}
library(dplyr)
library(arules)
library(arulesViz)
```

# 2.Set the working directory
```{r}
setwd("C:/Users/Ashley Rainbow/Desktop/Data Fellowship/data-fellowship/Association Rules/Association Rules Project/Association Rules project 2")
```

# 3.Import the data
```{r}
orders <- read.csv('Peak Association Rules.csv')
```

# 4.Check the data
```{r}
head(orders)
```
# 5.Change item name from a character string to factor.
```{r}
orders$Item.Name <- as.factor(orders$Item.Name)
```

# 6.Change item name from factor to character
```{r}
orders$Order.Number <- as.character(orders$Order.Number)
```

# 7.Group products by order number
```{r}
order_baskets = orders %>% 
  group_by(Order.Number) %>% summarise(order_basket = as.vector(list(Item.Name)))
```

# 8.View a sample order_basket
```{r}
head(order_baskets)
```

# 9.Convert to sparse matrix
```{r}
sparse_data <- as(order_baskets$order_basket, "transactions")
```

# 10.Inspect the sparse matrix
```{r}
summary(sparse_data)
inspect(sparse_data[1:20])
```

# 11.Plot the most popular products
```{r}
par(mar = c(4, 5, 1, 1))
itemFrequencyPlot(sparse_data, topN = 10, type = 'absolute', cex.names = 0.5, horiz = TRUE, xlim = c(0, max(1400)), col = c("#E64097"))
```

# 12.Create a rule for the most frequent itemset
```{r}
rules = apriori(data = sparse_data, parameter = list(target = "frequent itemsets",  support = 0.0004, minlen = 2), control = list(verbose = FALSE))
rules = sort(rules, by = 'support', decreasing = TRUE)
inspect(rules[1])
```

# 13.Create a dataframe base on rules data.
```{r}
order_rules_df <- DATAFRAME(sort(rules, by = "support", decreasing = FALSE))
```

# 14.Create a barplot from the orders dataframe.
```{r}
par(mar = c(5,17, 1, 1))
barplot(order_rules_df$support, main = "Order Rules", xlab = "Support", names.arg = order_rules_df$items, horiz = TRUE, las = 2, cex.names = 0.5, xlim=c(0, max = 0.001), col = c("#E64097"))
```

# 15.Finding the most common itemsets
```{r}
rules = apriori(data = sparse_data, parameter = list(support = 0.0004, confidence = 0.2, minlen = 2))
```

```{r}
summary(rules)
```

```{r}
inspect(rules[1:10])
```

# 16.Plotting most common sets of items being bought together.
```{r}
plot(rules, method = 'scatterplot', measure = c("support", "confidence"), shading = "lift", jitter = 1)
```

```{r}
plot(rules[1:10], method = 'graph', control = list(cex = 0.65))
```

# 17.Viewing rule sorted by sorted by confidence.
```{r}
inspect(sort(rules, by = 'confidence', decreasing = TRUE)[1:10])
```

# 18.Viewing rule sorted by sorted by support.
```{r}
inspect(sort(rules, by = 'support', decreasing = TRUE)[1:10])
```

# 19.Viewing the rule sorted by lift.
```{r}
inspect(sort(rules, by = 'lift', decreasing = TRUE)[1:10])
```

# 20.Finding the most common set of items bought with higher support but lower confidence.
```{r}
rules = apriori(data = sparse_data, parameter = list(support = 0.00055 , confidence = 0.1 , minlen = 2, maxlen = 3))
```

```{r}
summary(rules)
```

```{r}
inspect(rules)
```

# 21.Plotting most common sets of items being bought together with low confidence but high support.
```{r}
plot(rules, method = 'scatterplot', measure = c("support", "confidence"), shading = "lift", jitter = 1)
```

```{r}
plot(rules, method = "graph", control = list (cex = 0.65))
```

# 22.Creating Cream Toaster subset
```{r}
rules = apriori(data = sparse_data, parameter = list(support = 0.00003, confidence = 0.5, minlen = 2))
```

```{r}
cream_toaster_rule = sort(
  subset(rules, subset = rhs %in% 'Retro ST19020CN 4-Slice Toaster - Cream'),
  by = 'support', decreasing = T
)
inspect(cream_toaster_rule)
```

```{r}
lift_cream_toaster = sort(cream_toaster_rule, by = 'lift', decreasing = TRUE)
paste("Sorted by lift. ----------------------------------")
inspect(lift_cream_toaster)
```

```{r}
confidence_cream_toaster = sort(cream_toaster_rule, by = 'confidence', decreasing = TRUE)
paste("Sorted by confidence. ----------------------------------")
inspect(confidence_cream_toaster)
```

```{r}
plot(cream_toaster_rule, method = 'graph', control = list(cex = 0.7))
```


# 23.Generate summary reports for summary and lift rules.
```{r}
paste("Rules Report")
paste("Support")
summary(rules@quality$support)
paste("Lift")
summary(rules@quality$lift)
```

```{r}
paste("Cream Toaster Rules Report")
paste("Support")
summary(rules@quality$support)
paste("Lift")
summary(rules@quality$lift)
```

# 24.Comparing Cream Toaster rule results to the other rules.

```{r}
support_list <- rules@quality$support
mean_cream_toaster_support <- mean(cream_toaster_rule@quality$support)
lower_than_mean_rules <- support_list[support_list>c(mean_cream_toaster_support)]
support_rank <- length(lower_than_mean_rules)
support_rank
```
```{r}
paste("On average, the Cream Toaster has the", support_rank, "largest support of all the", length(rules), "rules")
```

```{r}
confidence_list <- rules@quality$confidence
mean_cream_toaster_confidence <- mean(cream_toaster_rule@quality$confidence)
lower_than_mean_rules_confidence <- confidence_list[confidence_list>c(mean_cream_toaster_confidence)]
confidence_rank <- length(lower_than_mean_rules_confidence)
confidence_rank
```
```{r}
paste("On average, the Cream Toaster has the", confidence_rank, "largest confidence of all the", length(rules), "rules")
```

```{r}
paste("In total", length(orders$Item.Name=="Retro ST19020CN 4-Slice Toaster - Cream"), "cream toasters were sold")
```

# Summary
As you can see above, the cream toaster is commonly bought with products of the same colour from the same brand. This suggests that selling these products as a bundle, for example a cream kitchen combination set, which would include a cream kettle and a cream bread bin could result in an increase in sales for products of this group.

>{Retro 4-Slice Toaster, Retro Jug Kettle, Retro Towel Pole, Retro Bread Bin Cream, Retro Microwave with Grill, Retro Traditional Kettle, Retro 45-litre Rectangular Bin, Solo Microwave, Retro 1-litre Canisters, Retro 2-piece Non-stick Frying Pan Set}